---
title: "ggmultiplot — 10 Software Style Gallery"
author: "ggmultiplot"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ggmultiplot Style Gallery}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 8, fig.height = 5)
library(ggmultiplot)
library(ggplot2)
```

## Overview

`ggmultiplot` lets you reproduce the default plot styles of 10 major
statistical graphing software packages with **a single function call**:
`ggchoice("style")`.

Each `ggchoice()` call applies three things at once:

1. A complete `theme_xxx()` — grid lines, borders, fonts, legend position
2. A `scale_color_xxx()` — discrete colour palette matching the software
3. A `scale_fill_xxx()` — discrete fill palette matching the software

**Scales are additive:** user-supplied `scale_*()` calls added *after*
`ggchoice()` take priority, so you can always fine-tune.

---

## Style Gallery

### 1. GraphPad Prism
*White background, no grid, black box border, sans-serif (Arial) font, bold title.
Prism pastel-to-bold colour defaults.*

```{r}
ggplot(mpg, aes(class, hwy)) +
  geom_boxplot_prism(aes(fill = class)) +
  ggchoice("prism") +
  labs(title = "GraphPad Prism Style")
```

### 2. SPSS (v12–24 default)
*Classic SPSS: white bg, light grey grid, full box border, sans-serif (Arial),
muted professional palette with SPSS blue (#3E58AC) dominant. Legend right,
title centred. Understated, publication-ready.*

```{r}
ggplot(mpg, aes(class, hwy)) +
  geom_boxplot_prism(aes(fill = class)) +
  ggchoice("spss") +
  labs(title = "SPSS Style")
```

### 3. OriginPro (classic default)
*Black-Red-Green-Blue palette, white bg, grid OFF (classic default),
black axes + layer frame, legend with box, bold title.*

```{r}
ggplot(mpg, aes(class, hwy)) +
  geom_boxplot_prism(aes(fill = class)) +
  ggchoice("origin") +
  labs(title = "OriginPro Style")
```

### 4. Stata (s2color scheme)
*Light bluish background tint (#F4F7FA), horizontal bluish grid only,
15-colour s2color palette, legend below plot. Default prior to Stata 18.*

```{r}
ggplot(mpg, aes(class, hwy)) +
  geom_boxplot_prism(aes(fill = class)) +
  ggchoice("stata") +
  labs(title = "Stata s2color Style")
```

### 5. Academic (AMS / Science)
*Minimal B&W, no grid, thin axis lines, bottom legend.
Ideal for journal submission.*

```{r}
ggplot(mpg, aes(class, hwy)) +
  geom_boxplot_prism(aes(fill = class)) +
  ggchoice("academic") +
  labs(title = "Academic Style")
```

### 6. SigmaPlot
*Single-series black, multi-series grayscale. White background,
no grid, black box frame. Publication-ready defaults.*

```{r}
ggplot(mpg, aes(class, hwy)) +
  geom_boxplot_prism(aes(fill = class)) +
  ggchoice("sigmaplot") +
  labs(title = "SigmaPlot Style")
```

### 7. JMP
*No grid, no outer frame (axes only), tick marks outside,
bottom legend. Clean interactive feel.*

```{r}
ggplot(mpg, aes(class, hwy)) +
  geom_boxplot_prism(aes(fill = class)) +
  ggchoice("jmp") +
  labs(title = "JMP Style")
```

### 8. MATLAB (R2014b+)
*Black outer box frame, 7-colour R2014b order (parula-like),
right legend. Anti-aliased modern look with inward ticks.*

```{r}
ggplot(mpg, aes(class, hwy)) +
  geom_boxplot_prism(aes(fill = class)) +
  ggchoice("matlab") +
  labs(title = "MATLAB Style")
```

### 9. Minitab
*Light grey background (#F5F5F5), dark blue frame (#1F497D),
no grid, black title. Dark blue strip headers.*

```{r}
ggplot(mpg, aes(class, hwy)) +
  geom_boxplot_prism(aes(fill = class)) +
  ggchoice("minitab") +
  labs(title = "Minitab Style")
```

### 10. MedCalc (ROC-optimised)
*Square plot area, clean white background, legend inside plot area
(lower right).  Optimised for ROC curve presentation; grid off for
uniform appearance in multi-panel layouts.*

```{r}
ggplot(mpg, aes(class, hwy)) +
  geom_boxplot_prism(aes(fill = class)) +
  ggchoice("medcalc") +
  labs(title = "MedCalc Style")
```

---

## Prism-style error bars

```{r}
df <- data.frame(
  group = c("Control", "Treatment A", "Treatment B"),
  mean  = c(10, 14, 8),
  sd    = c(2, 3, 1.5)
)

ggplot(df, aes(group, mean)) +
  geom_col_prism(fill = "#5B9BD5") +
  geom_errorbar_prism(aes(ymin = mean - sd, ymax = mean + sd)) +
  ggchoice("prism") +
  labs(title = "Prism T-bar Error Bars", x = "", y = "Value")
```

## Overriding scales

User scales added *after* `ggchoice()` take priority:

```{r}
ggplot(mpg, aes(class, hwy)) +
  geom_boxplot_prism(aes(fill = class)) +
  ggchoice("prism") +
  scale_fill_brewer(palette = "Set2") +   # overrides Prism fill
  labs(title = "Prism Theme + Custom Brewer Fill")
```

---

## Continuous Colour Scales

For heatmaps, surface plots, or any continuous-colour visualization, use
the `_c` suffix variants:

```{r}
# Simulated expression matrix
set.seed(42)
dm <- data.frame(
  gene  = rep(LETTERS[1:20], each = 10),
  sample = rep(1:10, 20),
  expr   = rnorm(200)
)

# MATLAB-style heatmap (Parula colormap)
ggplot(dm, aes(sample, gene, fill = expr)) +
  geom_tile() +
  ggchoice("matlab") +
  scale_fill_matlab_c() +
  labs(title = "MATLAB Parula Heatmap", x = "Sample", y = "Gene")

# Academic grayscale heatmap
ggplot(dm, aes(sample, gene, fill = expr)) +
  geom_tile() +
  ggchoice("academic") +
  scale_fill_academic_c() +
  labs(title = "Academic Grayscale Heatmap", x = "Sample", y = "Gene")
```

Available continuous scales: `scale_color/fill_prism_c()`,
`scale_color/fill_origin_c()`, `scale_color/fill_matlab_c()`,
`scale_color/fill_stata_c()`, `scale_color/fill_academic_c()`,
`scale_color/fill_minitab_c()`, `scale_color/fill_medcalc_c()`.

---

## Style Reference Table

| Software | Default Font | Grid | Outer Frame | Palette Type |
|----------|-------------|------|-------------|-------------|
| Prism | sans | Off | Black box (ticks outward) | Pastel-bold |
| SPSS | sans | Light grey | Black box | Blue-dominant |
| OriginPro | sans | Off | Black box (ticks outward) | Black→Red→Green→Blue |
| Stata s2color | sans | Horizontal bluish | Thin grey (legend bottom) | 15-colour cycle |
| Academic | sans | Off | Off (axes only) | Grayscale |
| SigmaPlot | sans | Off | Black box | Black + grayscale |
| JMP | sans | Off | Off (axes only) | JMP blue + Tableau |
| MATLAB | sans | Off | Black box | R2014b 7-colour |
| Minitab | sans | Off | Dark blue | Dark blue primary |
| MedCalc | sans | Off | Black box (legend inside) | Clinical contrast |

---

## Default ggplot2

Calling `ggchoice()` with no argument (or `ggchoice("ggplot2")`)
returns plain `theme_bw()` — no scale changes applied.
